---
title: "Explanatory analysis of a GEO dataset"
author: "Paula Balcells"
date: "May 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

What this exercise is about

# Summary

The study "Insights into responses of human astrocytes to H5N1 infection by transcriptional analysis""
was done using Homo Sapiens organisms and experiment type	Expression profiling by array.

The following description was provided:	
*Increasing evidence demonstrates influenza virus can not only affect the respiratory system, but also infect CNS and lead to CNS disorder and encephalopathy and encephalitis. Astrocytes are the most abundant cells in the CNS, which are capable of producing cytokines and neurotrophic factors and are essential for brain homeostasis and neuronal function. Previous studies suggested that influenza virus can infect astrocytes and induce proinflammatory cytokines response as well as apoptosis. Nevertheless, very few mechanistic data are available regarding host responses to H5N1 infection in astrocytes. In this study, a functional genomics approach was utilized to investigate comprehensive host responses to H5N1 infection in a human astrocyte cell line, U251 cells.*
 	
## Experimental design	
U251 cells were infected by H5N1 at MOI 1 or control . At time points 6, 12, and 24h, total RNA were extracted for microarray experiment. Three replicates were performed at each time point of infection and control infection.
 	
## Contributor(s)	
Lin X, Wang R, Sun X, Zhang J, Jin M

## Citation(s)	
Lin X, Wang R, Zhang J, Sun X et al. Insights into Human Astrocyte Response to H5N1 Infection by Microarray Analysis. Viruses 2015 May 22;7(5):2618-40. PMID: 26008703

# Data description

```{r readTargets}
targets <- read.csv("targets.txt", sep="")
expressions <- read.delim("GSE66597_expression.txt", row.names = 1)
```

```{r}
show(targets)
```

```{r}
dim(expressions)
summary(expressions)
```

```{r}
boxplot(expressions, col=targets$Color)
```